Change point analysis

### 2023-08-10 stats/supplements_changepoint_parameters.Rmd
adjust gaze model parts

```{r run_beasts, echo=T, message=FALSE, warning=FALSE, results=FALSE}
b1 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0), 
            tcp.minmax = c(0, 5),
            tseg.min = 10, 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)

b2 <- beast.irreg(y = y, time = x, season = "none", 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)

b3 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)

b4 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 10), 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)

b5 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 10), 
            tseg.min = 50, 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)

b6 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 10), 
            tseg.min = 100, 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)

b7 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 3), 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)

b8 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 20), 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)

b9 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 100), 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)

b10 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 100), 
            tseg.min = 100, 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)
```

@@

```{r actually_printbeasts}
beasts <- list(b1, b2, b3, b4, b5, b6, b7, b8, b9, b10)

counter = 1

@@

### Searching for reasonable minimum nr between change points

@@

mean_age_diff * mean_n_per_3y * 3

3/mean_age_diff
```

```{r final_beast, message=FALSE, warning=FALSE}
bfinal <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0), 
            tcp.minmax = c(0, 5),
            tseg.min = mean_age_diff * mean_n_per_3y * 3, 
            mcmc.chains = 4, mcmc.burnin = 1000, mcmc.samples = 10000, 
            ci = TRUE)
```

```{r final_beast}
showBeast(bfinal)

# pull out all relevant change points
cps <- tibble(
          cp = bfinal$trend$cp[1:bfinal$trend$ncp_median], 
          lowerCrI = bfinal$trend$cpCI[1:bfinal$trend$ncp_median,1], 
          upperCrI = bfinal$trend$cpCI[1:bfinal$trend$ncp_median,2]
       )
```

>>>>

```{r run_beasts, echo=T, message=FALSE, warning=FALSE, results=FALSE}
# this is the one we report in our paper: 
b0 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0), 
            tseg.min = 10, 
            mcmc.chains = 4, mcmc.burnin = 1000,  mcmc.samples = 10000, mcmc.seed = 42, 
            ci = TRUE)

b1 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0), 
            tcp.minmax = c(0, 5),
            tseg.min = 10, 
            mcmc.chains = 4, mcmc.burnin = 1000,  mcmc.samples = 10000, mcmc.seed = 42, 
            ci = TRUE)

b2 <- beast.irreg(y = y, time = x, season = "none", 
            mcmc.chains = 4, mcmc.burnin = 1000,  mcmc.samples = 10000, mcmc.seed = 42, 
            ci = TRUE)

b3 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            mcmc.chains = 4, mcmc.burnin = 1000,  mcmc.samples = 10000, mcmc.seed = 42, 
            ci = TRUE)

b4 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 10), 
            mcmc.chains = 4, mcmc.burnin = 1000,  mcmc.samples = 10000, mcmc.seed = 42, 
            ci = TRUE)

b5 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 10), 
            tseg.min = 50, 
            mcmc.chains = 4, mcmc.burnin = 1000,  mcmc.samples = 10000, mcmc.seed = 42, 
            ci = TRUE)

b6 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 10), 
            tseg.min = 100, 
            mcmc.chains = 4, mcmc.burnin = 1000,  mcmc.samples = 10000, mcmc.seed = 42, 
            ci = TRUE)

b7 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 3), 
            mcmc.chains = 4, mcmc.burnin = 1000,  mcmc.samples = 10000, mcmc.seed = 42, 
            ci = TRUE)

b8 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 20), 
            mcmc.chains = 4, mcmc.burnin = 1000,  mcmc.samples = 10000, mcmc.seed = 42, 
            ci = TRUE)

b9 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 100), 
            mcmc.chains = 4, mcmc.burnin = 1000,  mcmc.samples = 10000, mcmc.seed = 42,
            ci = TRUE)

b10 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 100), 
            tseg.min = 100, 
            mcmc.chains = 4, mcmc.burnin = 1000,  mcmc.samples = 10000, mcmc.seed = 42, 
            ci = TRUE)

b11 <- beast.irreg(y = y, time = x, season = "none", 
            torder.minmax = c(0, 0),
            tcp.minmax = c(0, 72), # as max nr, since 719 datapoints, 10 tseg.min
            tseg.min = 10, 
            mcmc.chains = 4, mcmc.burnin = 1000,  mcmc.samples = 10000, mcmc.seed = 42, 
            ci = TRUE)
```

@@

```{r actually_printbeasts}
beasts <- list(b0, b1, b2, b3, b4, b5, b6, b7, b8, b9, b10, b11)

counter = 0

@@

### To judge nr of data points between change points

@@

mean_age_diff
# [1] 0.11 corresponds to mean age difference between consecutive points on the x-axis

mean_n_per_3y
# [1] 27.65 data points correspond to 3 years of development
```

```{r}
b11

plot(b11)

str(b11)

# Y: a vector of length N; the estimated trend component.It is the Bayesian model averaging of all the individual sampled trend.
b11$trend$Y

plot(b11, vars = c('t', 'y'))

plot(b11$time, b11$trend$Y, type = 'l') # #directly plot output: the fitted trend

plot(b11$time, b11$trend$cpOccPr) # #directly plot output: trend chgpt occurrence prob
```

```{r}
est <- tibble(age = x, data = y, model = b11$trend$Y)
cps <- b11$trend$cp[1:b11$trend$ncp_median]

ggplot(data = est, aes(data, model)) + 
  geom_point() +
  geom_smooth(method = "lm") + 
  stat_cor() +
  scale_x_continuous(limits = c(0, 3), breaks = seq(0, 3, 0.5)) +
  scale_y_continuous(limits = c(0, 3), breaks = seq(0, 3, 0.5))

n_distinct(est$model)

est %>% 
  pivot_longer(cols = c(data, model), names_to = "source") %>% 
  ggplot(aes(x = age, y = value, color = as.factor(source))) + 
  geom_point(alpha = 0.5) +
  geom_vline(xintercept = cps, linetype = "dashed", alpha = 0.9, size = 0.75) +
  scale_color_manual(values = c("darkblue", "orange")) +
  scale_x_continuous(limits = c(3, 80), breaks = seq(0, 80, 5)) +
  scale_y_continuous(limits = c(0, 2.5), breaks = seq(0, 2.5, 0.5)) + 
  labs(x = "Age", y = "Mean imprecision", color = "Source")

est %>% 
  pivot_longer(cols = c(data, model), names_to = "source") %>% 
  filter(source == "model", age >= 20) %>%
  ggplot(aes(x = age, y = value)) + 
  geom_point() +
  geom_vline(xintercept = cps, linetype = "dashed", alpha = 0.9, size = 0.75) +
  scale_x_continuous(limits = c(20, 80), breaks = seq(0, 80, 5)) +
  scale_y_continuous(limits = c(0.5, 1), breaks = seq(0.5, 1, 0.5)) +
  labs(x = "Age", y = "Mean imprecision")
```
